### 2025-03-26 latex/backup_10_03_16h.tex

\documentclass[journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}

\begin{document}

\title{Deep transfer learning between Sentinel-1 acquisition modes enhances the segmentation of natural oil slick events in the Arctic}

\author{Julien Vadnais,~\IEEEmembership{Graduate Student Member,~IEEE, }Benjamin Aubrey Robson, Christian Haug Eide, Rune Mattingsdal, Malin Johansson
        % <-this % stops a space
\thanks{Manuscript received April 19, 2021; revised August 16, 2021. This research was carried out in the context of the InFluSe project (NFR Project No.?).}

%grouped affiliations
% \thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.
% Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.
% Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway. Corresponding author email address: julien.vadnais@uib.no}

%separate affilations
\thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.}
\thanks{Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.}
\thanks{Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}
\thanks{Corresponding author email address: julien.vadnais@uib.no. This article has supplementary available in the GitHub repository at $https:/github.com/julvad/temp_repo_a0$}

\thanks{}% <-this % stops a space
}

% The paper headers
\markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~14, No.~8, March~2025}%
{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

\IEEEpubid{\copyright This work is licensed under CC BY 4.0.}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

\maketitle

\begin{abstract}
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    The entirety of the data and Python code used in this study are provided in an open-access online repository.
\end{abstract}

\begin{IEEEkeywords}
    Article submission, IEEE, IEEEtran, journal, \LaTeX, paper, template, typesetting.
\end{IEEEkeywords}

\section{Introduction}
\IEEEPARstart{N}{atural} seepage is a significant contributor of oil entering marine environments \cite{kvenvoldenNaturalSeepageCrude2003}, yet the data required to update quantitative
estimates of its global scale remain insufficient, especially outside of the United States continental shelf \cite{nationalacademiesofsciencesengineeringandmedicineOilSeaIV2022}. Data scarcity, due to difficulties 
in monitoring remote offshore areas and obtaining continuous temporal observations, has limited the ability to study the ecosystem contributions and global behavior of natural hydrocarbon seepages. 
Spaceborne synthetic aperture radar (SAR) is well suited to this task due to its capacity to operate in any lighting conditions, its fine spatial resolution and its wide swath coverage. 
The capacity of SAR for detecting a slick relies on the damping ratio, the contrast between radar backscatter in slick and oil-free pixels 
\cite{hovlandSlickDetectionSAR1994,quigleyInvestigationDampingRatio2023}. By reducing sea surface roughness, which reduces radar backscatter, bodies of oil can be discernible from surrounding waters 
\cite{brekkeSAROilSpill2020,fingasReviewOilSpill2018,alpersOilsSurfactants2004}.

Numerous frameworks have been proposed to leverage SAR for automatizing slick detection using computer vision algorithms.
They face a common challenge which has been an ongoing topic of research over the past three decades: the presence of many oil slicks lookalikes including ocean eddies, biogenic slicks and newly-formed ice
\cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}. When exposed to a large number of labeled examples, 
deep learning algorithms, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or segmentation tasks \cite{goodfellowDeepLearning2016}. 
Although this approach can prove a powerful discriminator of slick lookalikes, it is known that when trained with limited sample sizes, models struggle to extract more abstract,
higher-level features \cite{bengioDeepLearnersBenefit2011,bengioDeepLearningRepresentations2012}. This issue is inherent to machine-learning-based SAR oil slick mapping, since slicks will only last 
a few hours on the surface \cite{jatiaultMonitoringNaturalOil2017,daneshgaraslHindcastModelingOil2017,oreillyDistributionMagnitudeVariability2022}, up to a day
\cite{macdonaldNaturalUnnaturalOil2015,macdonaldNaturalOilSpills1998}. Slick visibility is also dependent on a specific window of surface wind conditions, outside of which the sea either exhibits too many or 
too few capillary waves to produce a discriminative damping ratio \cite{quigleyInvestigationDampingRatio2023,sausDetectionDelineationProduced2021,gadeImagingBiogenicAnthropogenic1998}. 
\IEEEpubidadjcol  
Lastly, slick events are highly intermittent. Seepage rates vary considerably across both monthly and yearly timescales, with most active slicks having an occurence rate below 10\% 
\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}. Some seeps might stay dormant for several months or years before emitting oil again. In two different regions, it was found that only
half of the seeps show activity at any time \cite{jatiaultMonitoringNaturalOil2017,garcia-pinedaRemotesensingEvaluationGeophysical2010}, and it is not unsound to believe that this value might be even higher, 
given that dormant seeps are less likely to be discovered.

In this paper, we assess the potential of transfer learning for enhancing the automatic mapping of natural slicks. 
Deep transfer learning is a representation learning paradigm that aims to transfer valuable and generalizable features from one setting to another, where the contexts are similar but the target distributions differ 
\cite{goodfellowDeepLearning2016}.
We pre-trained two image segmentation models with oil slicks mapped in Sentinel-1 images over the North Sea, and then fine-tuned them on two known slick events in the 
Arctic, for which we have only limited observations. Over the Arctic, Sentinel-1 images are mostly acquired in the extra-wide (EW) mode with a dual HH/HV polarization. This differs from most previous studies 
on SAR oil slick mapping, which more often used the interferometric-wide (IW) mode acquired in the dual VV/VH polarization. Little work has until now addressed transfer learning between SAR acquisition modes, 
and natural oil slick events represent a prime example for a real-life application of deep learning segmentation with restricted labels.

\section{Data and methods}
\subsection{Data}
Evidence of two intermittent seep areas were recently documented in the Norwegian territorial waters. The first is located west of the Svalbard archipelago, offshore Prins Karls Forland (10.42°E, 78.49°N). 
It was revealed and thoroughly studied by Panieri et al. \cite{panieriArcticNaturalOil2024}. In 70 images in between 2015 and 2024, we manually delineated 81 oil slicks extending from a common seeping point.
The second area, located in the Barents Sea, was documented in \cite{serovWidespreadNaturalMethane2023} (but see also \cite{ivanovSearchDetectionNatural2020}). We mapped over 200 natural oil slicks from 88 images between 
2015 and 2020 over a spread-out area (31.33 - 32.37°E, 75.15°N - 75.35°N) in Sentralbanken high. This area is part of an active petroleum system with prospective hydrocarbon potential, though it remains largely 
underexplored \cite{lundschienNorthBarentsComposite2025}. Over those two seeps, labeled samples from Sentinel-1 EW images using the horizontal co-polarization (HH) band consistute the \textit{target domain}, $\mathit{D}_t$. 
At a lower latitude, oil slicks of various kinds can be observed in the North Sea, an hydrocarbon-rich and seepage-active region. \cite{juddSeabedFluidFlow2009,hovlandSeabedPockmarksSeepages1988}. 
With the same procedure as for the $\mathit{D}_t$, 1138 slicks were mapped in 361 images, but this time using the vertical co-polarization (VV) band of IW images. Here, labeled samples in the North Sea constitute the 
\textit{source domain}, $\mathit{D}_s$. Slicks numbers are given as an approximate, since often one or two slicks might have the same source, yet be mapped separately if their extent was not fully connected.

\subsection{Methods}
Our approach of transfer learning focuses on domain adaptation. In remote sensing, traditional domain adaptation methods imply that the $\mathit{D}_t$ and the $\mathit{D}_s$ have similar probability distributions 
between a common feature vector xi and a common label yi \cite{tuiaDomainAdaptationClassification2016}. In the case of deep learning (deep transfer learning) the

We trained two deep learning models: DeepLabV3 and the Segment Anything Model (SAM). DeepLabV3 is a fully convolutional neural network architecture for semantic segmentation which uses atrous convolutions 
and introduced atrous spatial pooling layers \cite{chenDeepLabSemanticImage2018}. On the other hand, SAM is a zero-shot learning vision transformer for segmentation \cite{kirillovSegmentAnything2023}. 
While the SAM itself does not assign classes, a variation implemented in the arcgis learn module extends its use to classification by trainable classification layers to the SAM's mask encoder 
\cite{esrideveloperFinetuneSegmentAnything2025}. Each model was trained for a maximum of 100 epochs, but using early-stopping, a commonly used way of limiting overfitting \cite{wangGeneralizingFewExamples2020}.
Two series of models were computed with two different backbones, since it was unknown whether a deep or a shallow model would perform best. For Deeplab, we used the ResNet-50 and ResNet-101 backbones, and for SAM, 
the ViT-B and Vit-L model checkpoints. ADD ABOUT DATA augm BATCH SIZE - DROPOUT ETC.

To calculate the gradients and update the parameters during training, a weighted binary cross-entropy loss function was used , defined by
\[
    L_\textit{CE} = -\frac{w}{N} \sum_{i=1}^{N} \left[ y_i \log(\hat{y}_i) + (1 - y_i) \log(1 - \hat{y}_i) \right]
\]

where \( w \) is the weight adjusting parameter, calculated as the inversely proportional scalar to the frequency of pixels per class \( w = \frac{P_c}{P} \) with \( P_C \) being the number of pixels 
of a given class \( C \).

Due to the particularly large pixel-number imbalance between the oil and water classes, we added a weighting term based on the segmentation Dice loss \cite{milletariVNetFullyConvolutional2016}:
% \begin{math}
\[
    L_\text{Dice} = 1 - \frac{2 \sum_{i=1}^{N} y_i \hat{y}_i}{\sum_{i=1}^{N} y_i + \sum_{i=1}^{N} \hat{y}_i}
\]
% \end{math}
This results in a simplified and version of the Combo loss from \cite{taghanakiComboLossHandling2019}, which gives the weighted loss function:
\[
    L_W = (1 - \alpha) \cdot L_\textit{CE} + \alpha \cdot L_\text{Dice}
\]
where $\alpha$ is a tunable hyperparameter in the form of a float number between 0 and 1, indicating the dice loss fraction.

Out of the 158 annotated Sentinel-1 EW HH scenes from the $\mathit{D}_t$, 128 were kept as a training split and 30 as a test split – about 19\%. Using bootstrapped sets from the training split, we trained
10 models for each power of 2 in the set \( N = \{ 2, 4, 8, 16, 32, 64\} \), where \( N \) is the number of training samples. Only one model was computed at \( N = 128\), using the full training set. For each model,
the Dice score was computed on the test split, allowing to calculate a 95\% statistical interval using the approach for classification recommended in \cite{burkovMachineLearningEngineering2020}.

Augmentation details: train deep learning page arcpy
Cross-entropy dice loss SaMED

\section{Results and Discussion}

Overfitting is the most important risk of training with very limited samples. 
Dropout
We find that freezing backbone weights improve. Converting to a 8bit raster with values rescaled on a 0,255 range provided similar performance with shorter inference time due to GPU processing larger batches. 
The smaller raster file format might also help with storage limitations.

\section{Conclusion}
Applications of transfer learning must be carefully evaluated, since in many cases its use can turn out inefficient, even detrimental if the context are too different \cite{mensinkFactorsInfluenceTransfer2022}.
In this study, we confirmed the potential of this approach by making use of labeled oil slicks in a different context to enhance the segmentation of natural oil slicks. We detailed pre-processing steps to
align Sentinel-1 SAR images captured in different acquisiton modes. These steps are crucial to ensure transferrable, high-level features between the source and the target domain.

Running two different deep learning models served not only for comparing purposes but also to extend our observations to a different model architecture. 
Convolutional nets can be rather limited do distribution shifts (domain change) \cite{hoffmanOneShotAdaptationSupervised2014,djolongaRobustnessTransferabilityConvolutional2021}
In both the convolutional neural network, DeepLabV3, and the vision transformer, SAM, 
yielded similar results.

% We found that very high dropout can improve model performance, especially when training with fewer samples. Considering the link between $\mathit{L_2}$-regularization and dropout \cite{wagerDropoutTrainingAdaptive2013}, 
% this finding is in line with Zhai et al. \cite{zhaiScalingVisionTransformers2022}, who found that a very strong L2-regularization reduced overfitting in the case of few-shot learning of a vision transformer.

This study learning between two different SAR acquisition.
A similar approach of two-stage fine-tuning was used for oil spill segmentation by Bianchi et al. \cite{bianchiLargeScaleDetectionCategorization2020a}, but instead downsizing resolution as the pretraining dataset.
Our best-performing model show comparable performance for natural oil slicks, needing only a small number of labeled samples, thus reducing computational costs and training time.

\section*{Acknowledgments}
Julien Vadnais received financial support from the Doctoral Research Scholarship of the Fonds de recherche du Québec.

{\appendix[Appendix A]
Write appendix or enter here
}

% \section{References}
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}